---
title: "Workfow to create image subsets for field mapping"
author: "Sarah Graves"
date: "Februrary 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

The purpose of this workflow is to have an automated way to create image subsets surrounding NEON plots. The basic process is to read in a spatial points file of all plot locations, identify the correct image file (either spatially or by reading an attribute in the point file), extract data and clip image to specified extent around plot, and save raster files to load onto the iPad.

Working with the specral data requires reading HDF5 files. It is necessary to install an R HDF5 package. See this [NEON website] (http://neondataskills.org/HDF5/Intro-To-HDF5-In-R/) for information for installing rhdf5 via bioconductor.

The functions in this workflow come from multiple packages and custom functions in a separate R file in the scripts folder.

```{r load-packages}

library(sp)
library(rgdal)
library(raster)
library(rhdf5)
library(plyr)

# install package from GitHub
#install_github("lwasser/neon-aop-package")
library(neonAOP)

# read in custom function
source("scripts/1.image_subsets/functions.R")

```


## Function overview

This step identifies the path number for a specified plot ID. The fight path was manually added to the spatial polygon shapefile. This was done because often there are two flight paths that intersect a single NEON plot center. Manual assignment of a NEON plot to a flight path ensures that the best flight path is chosen. The primary considerations for which path is optimal are 1) lack of clouds or cloud shadows, and 2) position in the center of the flight path to reduce look angle effects.

```{r Load-spatial-polygons}

find_flight_path

```

The next step is to find the HDF5 file that corresponds to the identified flight path. Often the files are stored in a single folder for the year and site. Therefore, the function should search through the list of file and find the file with the flight path in the name

```{r Identify-H5-file-path}

create_file_path


```

Once the specific file has been identified more complex functions to extract information from the HDF5 files can be used.

```{r HDF5-extraction}

find_point_coordinates

create_extent_around_point

```

## Plot parameters

These steps create objects that are used in all the functions below and are not specific to an image type. 

Each image subset is created around a spatial point data file. In this example, the point file is an ESRI Shapefile and is read in using the *rgdal* package.

The *find_point_coordinates* function requires the spatial points data frame, the name of the plot to be used, and the column where this attribute is located. If not using a external spatial points file, coordinates can be created from another source and used in the following step.

The *create_extent_around_point* creates a buffer of a specified distance in meters. The result is a square extent that has a length and width of the buffer distance * 2.


```{r specify-plot}

# read in shapefile
p_full <- readOGR("../data/NEON_plot_spatial_data/NEON plot shapefiles","OSBS_diversity_plot_centroids",
                  verbose = F,
                  stringsAsFactors=F)


plot_name <- "OSBS_017"


# 3. extract plot coordinates
# column 3 for tower file
# column 2 for diversity file
plot_coords <- find_point_coordinates(p_full,plot_name,2)

# 4. create extent around point
plot_extent <- create_extent_around_point(plot_coords,40)


```


## Create hyperspectral clips

These functions can be applied sequentially. The full workflow requires use of the neonAOP package by Leah Wasser.


```{r apply-functions}

# often data is stored in an external drive, need a pointer to this folder
#file_folder <- "D:/NEON_AOP_downloads/OSBS_Reflectance"

file_folder <- "F:/D03/OSBS/2014/OSBS_L1/OSBS_Spectrometer/Reflectance"

# folder to save images
# leave backslash on folder name
save_folder <-"../data/NEON_plot_spatial_data/digitizing_images/hs_fullBands"


# organgish image
#img_bands <- c(177,53,16)

# img_bands <- c(96,177,34)

#img_bands <- c(16,34,53,96,136,177)

name_bands <- paste(img_bands[1],img_bands[2],img_bands[3],sep="-")
img_file_name <- paste(plot_name,name_bands,sep="_")
#img_file_name <- paste(plot_name,"6bands",sep="_")

# 1. identify path
# path is column 19 for tower file
path <- find_flight_path(p_full,plot_name,plot_column=3,
                         path_column=19)

# 2. create path name
h5_file <- create_file_path(file_folder,path)

# 5. create extent object for H5 file
h5_extent <- create_extent(h5_file)

# create new extent object that can be applied in the create_stack function
extent_to_clip <- calculate_index_extent(clipExtent = plot_extent,
                                         h5Extent = h5_extent)

# 6. Create multi-band raster object
rgbimage <- create_stack(h5_file, 
             bands=img_bands,
             epsg = 32617,
             subset=T,
             dims = extent_to_clip)

plotRGB(rgbimage,stretch="lin")

# 7. Save raster to file
save_raster_image(raster_object = rgbimage,
                  image_type = "spectral",
                  save_folder = save_folder,
                  file_name = img_file_name, 
                  flatten = T)

```

# High res RGB camera data

## Function overview

The camera and lidar images are tiled on thousand meter tiles. The file names represent meters in the x and y direction.
To find which plot corresponds to which file, look up the x and y meters and round to the nearest thousands.

For x, it is XXX,000
For y is it YYYY,000


## Create RGB clip

```{r create-rgb-image}

save_folder <- "../data/NEON_plot_spatial_data/mapping_images/high_res/"
img_name <- paste("RGB",plot_name,sep="_")
source_folder <- "D:/D03/OSBS/2014/OSBS_L3/OSBS_Camera/"

rgb_clip <- create_RGB_clip(plot_coords = plot_coords,
                            source_folder = source_folder)

plotRGB(rgb_clip)

# use custom function to save the image
save_raster_image(rgb_clip,
                  image_type="camera",
                  save_folder = save_folder,
                  file_name = img_name,
                  flatten = T,
                  img_res = 0.25)

```


# Lidar data

The lidar-derived models of ground and surface elevation, and canopy height are available as raster files in the same tile system as the RGB images. The process is similar to loading RGB data and creating a subset, but only for a single band raster.

**NEXT** modify function to plot single band flattened image

```{r create-lidar-image}

save_folder <- "../data/NEON_plot_spatial_data/mapping_images/chm/"
img_name <- paste("CHM",plot_name,sep="_")
source_folder <- "D:/D03/OSBS/2014/OSBS_L3/OSBS_Lidar/CHM/"

chm_clip <- create_lidar_clip(plot_coords = plot_coords,
                            source_folder = source_folder,
                            image_type = "CHM")

plot(chm_clip)

# use custom function to save the image
save_raster_image(chm_clip,
                  image_type = "lidar",
                  save_folder = save_folder,
                  file_name = img_name,
                  flatten = T,
                  img_res = 1)

```



